---
title: "Lab week 5 - time series & a map"
author: "Adam Garber"
date: "2/6/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```


```{r}
library(tidyverse)
library(janitor)
library(lubridate)
library(here)
library(paletteer)
library(tsibble)
library(fable)
library(fabletools)
library(feasts)
library(forecast)
library(sf)
library(tmap)
library(mapview)

```


## Monthly US renewables
```{r}

us_renew <- read_csv(here("data", "renewables_cons_prod.csv")) %>% 
  clean_names()

```

- Make variable levels for "description" all lowercaseuse
- only keep observations where "description" variable contains "consumption"
- remove any obs where description level has the string "total"
```{r}

renew_clean <- us_renew %>%
  mutate(description = str_to_lower(description)) %>% 
  filter(str_detect(description, pattern = "consumption")) %>% # str_detect is logical TRUE/FALSE
  filter(!str_detect(description, pattern = "total"))

```

```{r}

renew_date <- renew_clean %>%
  mutate(yr_mo_day = lubridate::parse_date_time(yyyymm, "ym")) %>%
  mutate(month_sep = yearmonth(yr_mo_day)) %>% 
  mutate(value = as.numeric(value)) %>% 
  drop_na(month_sep, value)

class(renew_date$month_sep)

# parse data using lubridate::month()
renew_parsed <- renew_date %>% 
  mutate(month = month(yr_mo_day, label = TRUE)) %>% 
  mutate(year = year(yr_mo_day))
```

### Look at it:

```{r}

renew_gg <- ggplot(data = renew_date,
                   aes(x = month_sep,
                       y = value,
                       group = description)) +
  geom_line(aes(color = description))

renew_gg

# View(palettes_d_names) # to view all the palletes in paletteer
```

```{r}

renew_gg +
  scale_color_paletteer_d("ggsci::default_aaas")

```

### Coerce renew_parsed to a tsibble

```{r}

renew_ts <-  as_tsibble(renew_parsed,
                       key = description, index = month_sep)

```

```{r}

renew_ts %>% autoplot(value)
renew_ts %>%  gg_subseries(value)
renew_ts %>%  gg_season(value)

# to reproduce gg_season using ggplot
# ggplot(data = renew_parsed, aes(x=month, y=value, group = year)) +
#   geom_line(aes(color = year)) +
#   facet_wrap(~description, 
#              ncol = 1,
#              scales = "free",
#              strip.position = "right")
```

### just look at the hydroelectric energy consumption
```{r}
hydro_ts <- renew_ts %>% 
  filter(description == "hydroelectric power consumption")

hydro_ts %>% autoplot(value) 
hydro_ts %>% gg_subseries(value)
hydro_ts %>% gg_season(value)
```

### quarterly average consumption for hydro

```{r}

hydro_quarterly <- hydro_ts %>% 
  index_by(year_qu = ~(yearquarter(.))) %>% 
  summarise(avg_consumption = mean(value))

head(hydro_quarterly)

```

### decompose that hydro_ts

```{r}

dcmp <- hydro_ts %>% 
  model(STL(value ~ season(window = 5)))

components(dcmp) %>% autoplot()

# residual is ussually < 10% of original scale value
```


```{r}
hist(components(dcmp)$remainder)
```

```{r}
hydro_ts %>% 
  ACF(value) %>% 
  autoplot()
```

### warning, do more research

```{r}

hydro_model <- hydro_ts %>% 
  model(
    ARIMA(value)
  ) %>% 
  fabletools::forecast(h = "4 years")

hydro_model %>% autoplot(filter(hydro_ts, year(month_sep) > 2010))
```

### make a worldmap in a coupe of minutes

```{r}
world <- read_sf(dsn = here("data", "TM_WORLD_BORDERS_SIMPL-0.3-1"),
                 layer = "TM_WORLD_BORDERS_SIMPL-0.3")

mapview(world)
```






